SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-12 docs/questioning.tex
updated Figure 2

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For every arrow there is a
corresponding tuple that includes the transition number for
every attack. Every number indicates the number of the
steps taken until we reach this transition.
For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Most importantly,
Figure~\ref{fig:attacks} illustrates in which transitions
the defenses detect corresponding attacks.

A malicious user can initiate the attack through two different routes.
The attacker may use the browser of an innocent user as an attack
vehicle, through which the code will be injected in the application.
For example, the attacker could craft a malicious script into a {\sc url}
and then trick an end user to click on it via social engineering (i.e., phishing ---
{\it transition}: {\sc p-xss} 1.1 \text{\textbar} {\sc n-xss} 1.1 \text{\textbar} {\sc csrf} 1).
Alternatively, the attacker may be able to inject
directly the malicious code in the application. For example, consider
a web application that accepts and redisplays user input
without appropriate validation. An attacker could
upload data containing a specially crafted script (for example, in a
blog comment) to steal the cookies of the visiting users or embed
malicious {\sc sql} code to retrieve all the password entries from a
database ({\it transition}: {\sc dsl} 1 \text{\textbar}
{\sc p-xss} 1 \text{\textbar} {\sc n-xss} 1).
Note that an {\sc xss} attack can start from both routes.

@@

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request ({\it transition}: {\sc p-xss} 7 \text{\textbar} {\sc n-xss} 2).
Thus, in a non-persistent {\sc xss} attack the injected code
is never saved on the server and immediately becomes part of the
content that is sent back to a web user.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes a transition number for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc s}t{\sc b}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc e}i{\sc q}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point {\sc sdb}~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

Detection mechanisms vary on the location where they detect an attack.
There are four different points where an attack can be caught, as seen
in Figure~\ref{fig:attacks}:
\begin{itemize}
\item At the user's browser (Point {\sc ub}).
\item En route from the server to the user's browser --- in most cases
within a proxy (Point {\sc s}t{\sc b}).
\item Embedded in a query, but before it reaches the server's database
  (Point {\sc e}i{\sc q}).
\item After the malicious code has been executed on the server's
  database (Point {\sc sdb}).
\end{itemize}

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks, based on source code
injectiion. For each approach we provide the references
that present corresponding mechanisms:
{\bf 1}~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08,LV09},
{\bf 2}~\cite{BK04,JB07,GC09,APKLM10},
{\bf 3}~\cite{BWS05,SW06},
{\bf 4}~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\bf 5}~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\bf 6}~\cite{BV08,LV09,SMS13}.}
\end{figure*}

@@

{\it {\sc sql}rand}~\cite{BK04} implements {\sc isr} to detect {\sc
  sql} injection attacks. It allows programmers to create {\sc sql}
statements using randomized instructions instead of standard keywords.
The modified queries are reconstructed at runtime using the same key
that is inaccessible to the malicious user. {\sc sql}rand is one of
the few mechanisms that detect {\sc sql} injection attacks on the
database level (Point {\sc sdb}).

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
on the server-side, all the mechanisms that detect
such attacks are placed in some point within the server.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 (23\%) do so at the level
of the {\sc rdbms}---Point {\sc sdb}.
The other 11 wrap {\sc api} calls related to output
vectors---Point {\sc e}i{\sc q}.

@@

>>>>

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For every arrow there are
corresponding transition numbers for every attack.
Every number indicates the number of the
steps taken until we reach this transition.
For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Most importantly,
Figure~\ref{fig:attacks} illustrates in which transitions
the defenses detect corresponding attacks.

A malicious user can initiate the attack through two different routes.
The attacker may use the browser of an innocent user as an attack
vehicle, through which the code will be injected in the application.
For example, the attacker could craft a malicious script into a {\sc url}
and then trick an end user to click on it via social engineering (i.e., phishing ---
{\it transitions}: {\sc p-xss} 1.1, {\sc n-xss} 1.1, {\sc csrf} 1).
Alternatively, the attacker may be able to inject
directly the malicious code in the application. For example, consider
a web application that accepts and redisplays user input
without appropriate validation. An attacker could
upload data containing a specially crafted script (for example, in a
blog comment) to steal the cookies of the visiting users or embed
malicious {\sc sql} code to retrieve all the password entries from a
database ({\it transitions}: {\sc dsl} 1,
{\sc p-xss} 1, {\sc n-xss} 1).
Note that an {\sc xss} attack can start from both routes.

@@

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request ({\it transitions}: {\sc p-xss} 7, {\sc n-xss} 2).
Thus, in a non-persistent {\sc xss} attack the injected code
is never saved on the server and immediately becomes part of the
content that is sent back to a web user.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps-new.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there are
corresponding transition numbers for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub} (at the user's browser):~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc s}t{\sc b} (en route from the server to the browser):~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc e}i{\sc q} (embedded in a query en route to the database):~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point i{\sc db} (in the database):~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

Detection mechanisms vary on the location where they detect an attack.
There are four different points where an attack can be caught, as seen
in Figure~\ref{fig:attacks}:
\begin{itemize}
\item At the user's browser (Point {\sc ub}).
\item En route from the server to the user's browser --- in most cases
within a proxy (Point {\sc s}t{\sc b}).
\item Embedded in a query, but before it reaches the server's database
  (Point {\sc e}i{\sc q}).
\item After the malicious code reaches the server's database
  (Point i{\sc db}).
\end{itemize}

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks, based on source code
injectiion. For each approach we provide the references
that present corresponding mechanisms:
{\it Policy Enforcement}:~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08,LV09},
{\sc isr}:~\cite{BK04,JB07,GC09,APKLM10},
{\it Parse-Tree Validation}:~\cite{BWS05,SW06},
{\it Taint Tracking}:~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\it Training}:~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\it Hybrid}:~\cite{BV08,LV09,SMS13}.}
\end{figure*}

@@

{\it {\sc sql}rand}~\cite{BK04} implements {\sc isr} to detect {\sc
  sql} injection attacks. It allows programmers to create {\sc sql}
statements using randomized instructions instead of standard keywords.
The modified queries are reconstructed at runtime using the same key
that is inaccessible to the malicious user. {\sc sql}rand is one of
the few mechanisms that detect {\sc sql} injection attacks on the
database level (Point i{\sc db}).

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
on the server-side, all the mechanisms that detect
such attacks are placed in some point within the server.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 (23\%) do so at the level
of the {\sc rdbms}---Point i{\sc db}.
The other 11 wrap {\sc api} calls related to output
vectors---Point {\sc e}i{\sc q}.

@@

In this paper we first introduced an exploitation model
that can be a reference point which can be used by
researchers who wish to develop new countermeasures.
